Development and assessment of automated forest road projection methods using performance metrics relevant for wildlife

### 2024-06-10 analysis/paper/paper.Rmd
WIP updating results based on most recent run.

According to all aspatial performance measures, the grade penalty method for setting edge weights yields better results than our simple cost methods (compare Figure \@ref(fig:aspat-perf-fig) and Figure B.1). We therefore recommend using the grade penalty method, and focus on further examining variation in performance among grade penalty method variants (Figure \@ref(fig:aspat-perf-fig)). The MST regular high density method yields the best projections of road density within cutblocks, and all MST variants yield comparably good projections outside cutblocks, so overall the MST regular high density variant most accurately projects road density. All MST variants yield better projections of the forestry disturbance footprint than the Hardy or ILCP methods, and the Hardy and MST regular high density methods yield comparably accurate projections of road presence and distance to road.

@@

All of the road projection methods underestimate road density, road presence and forestry disturbance footprint (Figure \@ref(fig:aspat-perf-fig)). In other words, these road network projection methods project fewer roads than we observe. Projections of distance to road, the forestry disturbance footprint, and the road disturbance footprint were more accurate than projections of road density and road presence (Figure \@ref(fig:aspat-perf-fig)), suggesting that these measures were more robust to our coarse assumptions about road projection. One reason for underestimation of road density is that all these methods generate straight road segments within a pixel. As raster resolution increases, the discrepancy within pixels between real roads and straight projected segments increases, which explains why projections of road density are less accurate when the raster resolution is coarse (Figure D.1).

@@

Although the MST method performed better than the alternatives we tested in terms of several important wildlife habitat metrics (Figure \@ref(fig:aspat-perf-fig)), it requires greater computation resources for large landscapes (Figure S7). All of these methods require construction of a large mathematical graph with a node for each pixel in the weight raster where road construction is possible (i.e. not NA), and as such, memory requirements will increase with the number of nodes in the graph (Figure S7). Compute time also increases with the number of landing targets (Figure S7). Lochhead et al's [-@lochhead_demo_2018; -@lochhead_linking_2022] method of graph construction is more computationally efficient than the adjacency matrix approach used in the R package `gdistance` [@van_etten_r_2017] (Figure S8), in part because it avoids the additional cost of constructing a transition matrix, and in part because it ignores variation in the size of raster cells that is accounted for in the `gdistance` _geoCorrection_ function. The Hardy QGIS plugin is more than an order of magnitude faster than the MST and ILCP methods implemented in R (Figure S7).

>>>>

According to all aspatial performance measures, the grade penalty method for setting edge weights yields better results than our simple cost methods (compare Figure \@ref(fig:aspat-perf-fig) and Figure B.1). We therefore recommend using the grade penalty method, and focus on further examining variation in performance among grade penalty method variants (Figure \@ref(fig:aspat-perf-fig)). The MST regular high density method yields the best projections of road density overall, while the MST random high density method had the best projections within cutblocks, and all MST variants yield comparably good projections outside cutblocks. All MST variants yield better projections of the forestry disturbance footprint than the Hardy or ILCP methods, and the Hardy and MST regular high density methods yield comparably accurate projections of road presence and distance to road.

@@

All of the road projection methods underestimate, road presence and forestry disturbance footprint (Figure \@ref(fig:aspat-perf-fig)). In other words, these road network projection methods project fewer roads than we observe. Projections of distance to road, the forestry disturbance footprint, and the road disturbance footprint were more accurate than projections of road density and road presence (Figure \@ref(fig:aspat-perf-fig)), suggesting that these measures were more robust to our coarse assumptions about road projection. One reason for underestimation of road density is that all these methods generate straight road segments within a pixel. As raster resolution increases, the discrepancy within pixels between real roads and straight projected segments increases, which explains why projections of road density are less accurate when the raster resolution is coarse (Figure D.1).

@@

Although the MST method performed better than the alternatives we tested in terms of several important wildlife habitat metrics (Figure \@ref(fig:aspat-perf-fig)), it requires greater computation resources for large landscapes (Figure S7). All of these methods require construction of a large mathematical graph with a node for each pixel in the weight raster where road construction is possible (i.e. not NA), and as such, memory requirements will increase with the number of nodes in the graph (Figure S7). Compute time also increases with the number of landing targets (Figure S7). Lochhead et al's [-@lochhead_demo_2018; -@lochhead_linking_2022] method of graph construction is more computationally efficient than the adjacency matrix approach used in the R package `gdistance` [@van_etten_r_2017] (Figure S8), in part because it avoids the additional cost of constructing a transition matrix, and in part because it ignores variation in the size of raster cells that is accounted for in the `gdistance` _geoCorrection_ function. The Hardy QGIS plugin is 65 times faster than the MST and ILCP methods implemented in R (Figure S7).
